---
title: "Project4"
author: "Shahid Abdulaziz"
date: "11/29/2020"
output:
  word_document: default
  html_document: default
---

# Problem 1

**Read the dataset in Batteries_Life.csvPreview the document into R. Name the loaded data Batteries_Life. And use the head() function to show the first 6 rows.**

```{r, echo = TRUE}
library(readr)
library(stats)
library(ggplot2)
library(sqldf)



Batteries_Life <- read_csv("C:/Grad School/DSCI 618/Week4/Batteries_Life.csv") # importing data

head(Batteries_Life,6) # getting the first six rows of the data
```

# Problem 2

**Conduct the analysis of variance and test hypotheses on the main effects. What conclusions can be drawn using significance level=0.05?**

```{r, echo = TRUE}
aov1 = aov(Batteries_Life$Life ~Batteries_Life$Brand + Batteries_Life$Device ) #first anova test without interaction

summary(aov1) #anova results


```

At a 95% confidence level, we would fail to reject the null hypthoesis for Brand and their Devices because there is evidence to prove there is a difference in battery life for brands with a pe-value of 0.0112 and devices with a p-value of 4.5e-07.

# Problem 3

**Construct an interaction plot.**

```{r, echo = TRUE}
ggplot(Batteries_Life, #ggplot for interactions 
       aes(x=Batteries_Life$Device,y=Batteries_Life$Life,
           color=Batteries_Life$Brand,
           group=Batteries_Life$Brand,
           label=Batteries_Life$Life))+
    geom_line(size=.5)+
    geom_point(size=1)+
    geom_text(aes(label=Batteries_Life$Life),hjust=0, vjust=-0.5)+
    labs(title = "Experiment With Interaction",
         x = "Devices",
         y = "Battery Life")


```

# Problem 4 

**Test on the interaction between factors. What conclusions can be drawn using significance level=0.05?**

```{r, echo = TRUE}
aov2 = aov(Batteries_Life$Life ~Batteries_Life$Brand * Batteries_Life$Device ) #second anova test with interaction

summary(aov2) #anova results


```
At a 95% confidence level, we would fail to reject the null hypthoesis for the interaction  between Brand and their Devices because there is evidence to prove there is a difference in battery life a  p-value of 1.13e-05 .

# Problem 5

**Prepare appropriate residual plots and comment on the model’s adequacy.**

```{r, echo = TRUE}
plot(aov2) #plotting the risduals of the second anova test

```

## QQ Plot
The residuals are normally distributed  because they are around the line of best fit

## Residuals vs Fitted

The residuals  do not have a linear trend because they are spread out around the horizontal line

## Scale-Loaction

The residuals do have homoscedasticity because they are  equally distributed among the horizontal line.  



# Problem 6

**Which brand of batteries would you recommend?**

```{r, echo = TRUE}
sqldf("select  brand, avg(life) from Batteries_Life group by  brand " ) #looking at the avg 
```

I would suggest going with Brand B since the overall avg life of the batteries is higher than the Brand A of the batteries.

(I do not know if the question what looking for just the overall brand of the individual devices within the brands)
