---
title: "Project5_Abdulaziz_Shahid"
author: "Shahid Abdulaziz"
date: "12/6/2020"
output: word_document
---

# Problem 1

**Read the dataset in Yield.csv Preview the document into R. Name the loaded data Yield. And use the head() function to show the first 6 rows.**



```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(sqldf)
library(gridExtra)

Yield <- read_csv("C:/Grad School/DSCI 618/Week5/Yield.csv") #reading in my data


head(Yield, 6) #looking at the first six rows


str(Yield)

Yield$Blocks = as.factor(Yield$Blocks)
Yield$Temperature = as.factor(Yield$Temperature)


str(Yield)
```

# Problem 2

**Examine Data Graphically using strip charts, box plots, or other charts that can help you check if treatment and block factors affect the yield.**

```{r, echo = TRUE}
a = ggplot(Yield, aes(Blocks, Yield, fill = Temperature))+
  geom_boxplot()

b = ggplot(Yield, aes(Blocks, Yield, fill = as.factor(Pressure)))+
  geom_boxplot()+
  labs(fill = "Pressure")

grid.arrange(a,b )
```

# Problem 3

**Explain if the blocking factor is necessary for this experiment.**

```{r, echo = TRUE}
aov1 = aov(Yield ~., Yield) #anova with all variables
aov2 = aov(Yield ~.-Blocks, Yield) #anova without blocker
summary(aov1)
summary(aov2)
```
Blocking factor is necessary for this experiment with a p-value of  0.000196. Important to note that without the blocks, pressure is not longer significant with a p-value of 0.186.

# Problem 4

**Conduct the analysis of variance and test hypotheses on the main effects. What conclusions can be drawn using significance level=0.05.**

```{r, echo = TRUE}
summary(aov1)
```

With a 95% confidence level we can reject the null hyoithesis because there is evidence to show all variable have a significant influence on Yield.



# Problem 5

**Construct an interaction plot.**
```{r, echo = TRUE}


YieldDay1 = sqldf("Select * from Yield where blocks = 'Day1'")
YieldDay2 = sqldf("Select * from Yield where blocks = 'Day2'")

with(YieldDay1 , interaction.plot(Pressure,Temperature,Yield,type="b",col=10:13))
title(main = "Day One Blocker")

with(YieldDay2 , interaction.plot(Pressure,Temperature,Yield,type="b",col=10:13))
title(main = "Day Two Blocker")


```

# Problem 6

**Test on the interaction between factors. What conclusions can be drawn using significance level=0.05?**
```{r, echo = TRUE}
aov3 = aov(Yield ~.^3, Yield) #adding interaction with my variables
summary(aov3)

```

at a 95% confidence level, we fail to reject the null hypthosis because there is not enough evidence to support that there is an interaction amongst the factor variables.