---
title: "Homwork2_Abdulaziz_Shahid"
author: "Shahid Abdulaziz"
date: "1/19/2021"
output: word_document
---


```{r}
library(socviz)
library(ggplot2)
library(grid)


```

# Problem 1 
The image below (Favorability on Taking Office) describes the favorability of seven US presidents upon taking office. Improve the above image by creating a simple bar chart.


```{r}



df = data.frame(colvalues = c(79,78,68,65,58,56,40),
                presidents = c("Obama","Carter","Clinton","G.W Bush","Reagon","G.H.W Bush","Trump"),
                party = c("Democrat","Democrat","Democrat","Republican","Republican","Republican","Republican" ))

question1 =      ggplot(df, aes(reorder(as.factor(presidents),colvalues )))+
                  geom_bar(aes( y = colvalues, fill = party ),stat = "identity")+
                  labs(x ="Presidents",
                       y= "Favorability",
                       title = "Favorability on taking office")

ggsave("Question1_Abdulaziz_Shahid.png",question1)
question1 
```

# Problem 2
For the gss_sm dataset discussed in the lecture notes

## A 
First, replicate the image below but have the color of the proportions vary by religion.


```{r}


question2a =    ggplot(data = gss_sm, aes(x = religion,fill = factor(..x..) )) +
               geom_bar(aes(y = ..prop.. ,
                       group = 1),
                       stat = "count",
                       show.legend =  FALSE)
ggsave("Question2A_Abdulaziz_Shahid.png",question2a)
question2a
```

## B

After completing the replication.refer to the lecture notes where we plotted: ggplot(data = gss_sm, aes(x = bigregion)) + geom_bar(aes(fill = religion), position = "dodge"). Repeat this plot with proportions in the y-axis instead of count.

```{r}


question2b=    ggplot(data = gss_sm, aes(x = bigregion)) +
                 geom_bar(aes(y=..prop..,fill = religion, group = religion ), position = "dodge") 

ggsave("Question2B_Abdulaziz_Shahid.png",question2b)
question2b
```

#Problem 3
For the election dataset (this is part of the socviz package). This dataset contains State-level vote totals and shares for the 2016 US Presidential election. Replicate the image below.

```{r}


text_high <- textGrob("(Trump)", gp=gpar(fontsize=5, fontface="bold"))
text_low <- textGrob("(Clinton)", gp=gpar(fontsize=5, fontface="bold"))


question3 =    ggplot(election, aes(y = reorder(st,r_points) ,color = winner ))+
               geom_vline(xintercept = 0, linetype = "solid")+
               geom_point(aes(x =r_points),show.legend =  FALSE) +
               facet_wrap(~census,scales = "free_y", nrow = 2) + 
               annotation_custom(text_high,xmin=50,xmax=50,ymin= -2,ymax=-1.5) +
               annotation_custom(text_low,xmin=-30,xmax=-30,ymin=-2,ymax=-1.5)+
               scale_x_continuous(limits=c(-35,50),labels = abs, n.breaks = 10) +
               coord_cartesian(clip = "off")+
               scale_color_manual(values=c("#2E74C0", "#CB454A"))+
               labs(
                 x = "Point Margin",
                 title = "US Presidential Election 2016, State-level results ",
                 y = "")



ggsave("Question3_Abdulaziz_Shahid.png", question3)
question3

```