---
title: "Project 5"
author: "Shahid Abdulaziz"
date: "7/31/2020"
output: word_document
---


# Problem 1

**Load the dataset bike.csvPreview the document into memory. Then split the data into a training set containing 2/3 of the original data (test set containing remaining 1/3 of the original data).**

```{r , echo = TRUE, warning = FALSE}
library(readr)
library(caret)
library(tree)
library(randomForest)

Bike  <- read_csv("C:/Grad School/DSCI 512/Week 5/Data/Bike (1).csv", #Reading in the data and making sure the data class types are right
    col_types = cols(season = col_factor(levels = c("1", 
        "2", "3", "4")), holiday = col_factor(levels = c("0", 
        "1")), weather = col_factor(levels = c("1", 
        "2", "3", "4"))))


Bike_Part <- createDataPartition(Bike$count, p = .66, list = FALSE, times =1) #partitioning my data for testing and trianing
Bike_Train <- Bike[Bike_Part,] #Creating training data set
Bike_Test <- Bike[-Bike_Part,] #Creating testing data set

```

# Problem 2

**Build a tree model using function tree().**

*A:* The response is count and the predictors are season, holiday, workingday, temp, atemp, humidity, windspeed, casual, and registered.

*B:* Perform cross-validation to choose the best tree by calling cv.tree().

*C:* Plot the model results of b) and determine the best size of the optimal tree.

*D:* Prune the tree by calling prune.tree() function with the best size found in c).

*E:* Plot the best tree model.

*F:* Compute the test error using the test data set

**A:**

```{r , echo = TRUE, warning = FALSE}
set.seed(19981) #Creating a seed

Tree_Model_A <- tree(count~season+ holiday+ workingday+temp+atemp+humidity+windspeed+casual+registered, data = Bike_Train) #Creating a tree model to predict count
```

**B:**

```{r , echo = TRUE, warning = FALSE}
Best_Tree_Model_A <- cv.tree(Tree_Model_A ) #Performing cross-validation for the best tree
```

**C:**

```{r ,echo = TRUE, warning = FALSE}
plot(Best_Tree_Model_A$size,Best_Tree_Model_A$dev, ttpe = 'b') #Plotting the modle results
```

5 - 8 Model sizes are roughly the same so I will pick Size 5 model as the best one because it is a simp models than the others with less nodes. 

**D:**
```{r , echo = TRUE, warning = FALSE}
prune_Best_Tree_Model_A <- prune.tree(Tree_Model_A , best = 5) #pruning the tree whihc I thought was the best
```

**E:**
```{r , echo = TRUE, warning = FALSE}
plot(prune_Best_Tree_Model_A) #plotting the best tree
text(prune_Best_Tree_Model_A, pretty = 0)
```

**F:**
```{r , echo = TRUE, warning = FALSE}

Predict_Prune_Best_Tree_Model_A <- predict(prune_Best_Tree_Model_A, data = Bike_test) #predicting reuslts with mt testing dataset

mean((Predict_Prune_Best_Tree_Model_A - Bike_Test$count )^2) #Calculating MSE for my best tree model

```

# Problem 3

**Build a random forest model using function randomForest()**

**A:**   The response is count and the predictors are season, holiday, workingday, temp, atemp, humidity, windspeed, casual, and registered.

**B:**   Compute the test error using the test data set.

**C:**  Extract variable importance measure using importance() function.

**D:**   Plot the variable importance using function varImpPlot(). Which are the top 2 important predictors in this model?

**A:**
```{r , echo = TRUE, warning = FALSE}
Random_Forest_Model <- randomForest(count~season+ holiday+ workingday+temp+atemp+humidity+windspeed+casual+registered, data = Bike_Train, importance = TRUE) #Creating my random forest model
```

**B:**

```{r , echo = TRUE, warning = FALSE}
Random_Forest_Model_Prediction <- predict(Random_Forest_Model, newdata= Bike_Test) #Creating a random forest model

mean((Random_Forest_Model_Prediction- Bike_Test$count )^2) #Calculating MSE
```

**C:**

```{r , echo = TRUE, warning = FALSE}
importance(Random_Forest_Model ) #looking at the variable importance
```

**D:**

```{r , echo = TRUE, warning = FALSE}
varImpPlot(Random_Forest_Model) #Plotting my variable importance
```

Registered and casual are the top two most important variables for MSE and Node Purity. 
