---
title: "Assignment 5"
author: "Shahid"
date: "11/8/2019"
output: html_document
---



Special Note: This assigment is something I created that will be given to a college 300 level business course. This course is an introduction to business data analytics using R. This is the answer key below.




Your manager wants the last four weeks of sales data from the PA region. Additionally, your manager wants to see three aggregations.
Aggregations Wanted:
1:  The Entire amount of sales for the city of Harrisburg in date order. 
2: The total sales for each product by every city
3: The entire sales for city in date order except for the city of Lancaster. 
 The data however cannot fit into an excel spreadsheet because the rows are passed the maximum amount allowed in Excel. Additionally, each week of data is outputted in a separate CSV file. You decide using R would be most appropriate in combining the datasets.  The outline of your tasks is below, make sure to complete each one.

1:   Develop a method to read the csv files from a folder into a list in R.
 
2: Take the list of files within R and combine them together into one data frame. Note, there are many different methods to do this.

3: Ensure the files have combined successfully. 

4:  Clean the Data
 
5:  Perform aggregations 

Students, remember that there is always more than one way to skin a cat. 

```{r, message= FALSE}
library(gtools)
library(lubridate)
library(tidyverse)
library(dplyr)
library(sqldf)
library(plyr)

setwd("C:/Users/shahi/Desktop/Sanjay/Assignment5")
FileList <- list.files(path = choose.dir("C:/Users/shahi/Desktop/Sanjay/Assignment5"), pattern = "*.csv")

FileList


FL <- lapply(FileList, read.csv)

FL


lapply(FileList, cbind, ID = c(" "))

newlist <- Map(cbind, FL, ID = (1:length(FL)))

Combined <- do.call(smartbind, newlist)

Combined


 Rename <-function(x,ColNames){
      names(x) <- ColNames
      return(x)
 }
 
 
Colnames1 <- c("Date", "City", "Product", "Sales", "ID")
ColnamesFixed <- lapply(newlist ,  Rename , Colnames1) 
ColnamesFixed 

EntireCombine <- do.call(rbind, ColnamesFixed )

EntireCombine 
str(EntireCombine )

as.data.frame(EntireCombine)

EntireCombine$Date <- as.Date(EntireCombine$Date)

parse <- parse_number(EntireCombine$Sales)


EntireCombine$Sales <- sapply(EntireCombine$Sales, parse_number)

EntireCombine$City <- as.character(EntireCombine$City)
str(EntireCombine)


EntireCombine$City <- revalue(EntireCombine$City, c("Harrisburgh "="Harrisburg"))
EntireCombine$City <- revalue(EntireCombine$City, c("Pittsburgh "="Pittsburg"))

EntireCombine

Harrisburg_Only <- sqldf("SELECT Date, City, sum(Sales) AS Sales
       FROM EntireCombine
       WHERE City LIKE 'Harris%'
       GROUP BY Date, City
      ")

Harrisburg_Only 


Harrisburg_Only_By_Product <- sqldf("SELECT Date, City, sum(Sales) AS Sales,Product
       FROM EntireCombine
       WHERE City LIKE 'Harris%'
       GROUP BY Date, City, Product
      ")
Harrisburg_Only_By_Product


City_By_Product <- sqldf("
SELECT City, Product, sum(Sales) as Sales
FROM EntireCombine
       GROUP BY City, Product
       ORDER BY Sales ASC
      ")

City_By_Product 

By_Date_City <- sqldf("SELECT Date, City, sum(Sales) AS Sales
                        FROM EntireCombine
                        WHERE CITY NOT LIKE ('Lan%')
                        GROUP BY DATE, CITY
                      ")
By_Date_City
```


```{r}
library(ggplot2)



ggplot(Harrisburg_Only, aes(Date, Sales))+
   geom_line( colour = "red")+
   labs(title = "Harrisburg Sales",
        subtitle = "Jan 13th - Feb 3rd"
)

ggplot(Harrisburg_Only, aes(Date, Sales))+
   geom_bar(stat = "identity", position =  "dodge")+
   labs(title = "Harrisburg Sales",
        subtitle = "Jan 13th - Feb 3rd"
)

ggplot(Harrisburg_Only_By_Product, aes(Date, Sales, fill = Product))+
   geom_bar(stat = "identity", position =  "dodge")+
   labs(title = "Harrisburg Sales By Product",
        subtitle = "Jan 13th - Feb 3rd"
)



ggplot(By_Date_City, aes(Date, Sales, fill = City))+ 
   geom_line(aes(colour = City))+
   labs(title = "All Cities Total Sales",
        subtitle = "Jan 13th - Feb 3rd"
        )


ggplot(By_Date_City, aes(Date, Sales, fill = City)) +
   geom_bar( stat = "identity", position =  "dodge")+
   labs(title = "All Cities Total Sales",
        subtitle = "Jan 13th - Feb 3rd"
        )

ggplot(EntireCombine, aes(Date, Sales, fill = Product ))+
      geom_bar( stat = "identity", position =  "dodge")+
   labs(title = "All Cities Total Sales By Product",
        subtitle = "Jan 13th - Feb 3rd"
        )






````
