---
title: "Project 6"
author: "Shahid Abdulaziz"
date: "8/8/2020"
output:
  word_document: default
---

# Problem 1

**Load the dataset bike.csvPreview the document into memory. Convert holiday to a factor using factor() function. Then split the data into training set containing 2/3 of the original data (test set containing remaining 1/3 of the original data).**



```{r , echo = TRUE, warning = FALSE}

library(readr)
library(caret)
library(e1071)

Bike  <- read_csv("C:/Grad School/DSCI 512/Week 6/Data/Bike (1).csv", #Reading in the data and making sure the data class types are right
    col_types = cols(season = col_factor(levels = c("1", 
        "2", "3", "4")), holiday = col_factor(levels = c("0", 
        "1")), weather = col_factor(levels = c("1", 
        "2", "3", "4"))))


Bike_Part <- createDataPartition(Bike$count, p = .75, list = FALSE, times =1) #partitioning my data for testing and training
Bike_Train <- Bike[Bike_Part,] #Creating training data set
Bike_Test <- Bike[-Bike_Part,] #Creating testing data set


```



# Problem 2

**2.** Build a support vector machine model.

  **a.**    The response is holiday and the predictors are: season, workingday, casual, and registered. Please use svm() function with radial kernel and gamma=10 and cost = 100.
```{r , echo = TRUE, warning = FALSE}
set.seed(989898) #creating a seed

SVM_Model1 <- svm(holiday ~ season + workingday + casual + registered, data = Bike_Train,  kernal = "radial", gamma = 1, cost = 100 ) #Creating my SVM model

SVM_Model1

```

  **b.**    Perform a grid search to find the best model with potential cost: 1, 10, 50, 100 and potential gamma: 1, 3, and 5 and using radial kernel and training data set.

```{r , echo = TRUE, warning = FALSE}

Best_SVM_Model <- tune(svm,holiday ~ season + workingday + casual + registered, data = Bike_Train,  kernal = "radial", gamma = c(1,3,5), ranges = list(cost = c(1,10,50,100))) #Selecting the best SVM model


```

  **c.**    Print out the model results. What’s the best model parameters?

```{r , echo = TRUE, warning = FALSE}

summary(Best_SVM_Model) #summary of the best svm model
```

Cost of 10 is the best paramter.

  **d.**    Forecast holiday using the test dataset and the best model found in c).

```{r , echo = TRUE, warning = FALSE}


SVM_Predictions <-predict(Best_SVM_Model$best.model, Bike_Test) # Predicting based off of the tets set

```

  **e.**    Get the true observations of holiday in the test data set.

```{r , echo = TRUE, warning = FALSE}

Holiday_TrueObv <- Bike_Test$holiday #Getting the true observations


```

  **f.** Compute the test error by constructing the confusion matrix. Is it a good model?

```{r , echo = TRUE, warning = FALSE}

confusionMatrix(Holiday_TrueObv,SVM_Predictions ) #creating a confusion matrix


```


With an accuracy of 97.61% the model is very good.

